The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-01-27 paper/sections/methods.tex
12 pages

\section{Physics-Based Animation}
Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
physics-based animation, specifically elastodynamics. 
The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\dave{cite Lanczos again} to arrive at 
the pertinent equations of motion:
\begin{equation*}
    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \fext,
\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
vector of internal forces and $\fext\in\real^{n}$ are external forces such due to gravity or contact. 
We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
Because SEM yields equations of motion in the standard form, it is compatible with any standard integration scheme, though in this submission we limit ourselves to
Implicit Euler.

@@

We derive the mass matrix for our system from the definition of kinetic energy:
\begin{equation}
    T = \dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
\end{equation}
% \begin{equation}
% T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
% \text{,}
% \end{equation}
% where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain. Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
% \begin{equation}
% \begin{split}
% T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
% \end{split}
% \text{.}
% \end{equation}
% The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.

@@

\subsection{External Forces}
We compute the force of gravity and apply external forces in the standard way --
$\vc{f}_g = \int_{\Omega} \rho\vc{g} d\Omega$, where $\vc{g}$ is the acceleration due to gravity acting on the object. 
We leverage our meshless quadrature~(\refsec{quadrature}) for integration.

@@

The total error at the boundary of the object can be written as
\begin{equation}
\epsilon = \int_{\partial\Omega}||\Gamma\left(\refX\right)\vc{q} - \Jnurbs\left(\refX\right)\vc{q}||_2^2 d\refX \\
\end{equation}, which is a quadratic.

>>>>

\section{Physics-Based Animation}
Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
physics-based animation, specifically elastodynamics. 
The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\cite{lanczos2012variational} to arrive at 
the pertinent equations of motion:
\begin{equation*}
    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \vc{f}_g + \fext,
\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\ddot{\vc{q}}\in\real^{n}$ are the generalized accelerations, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
vector of internal forces, $\vc{f}_g\in\real^{n}$ are body forces such as gravity and $\fext\in\real^{n}$ are external forces such as those due to contact. 
We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
Because SEM yields equations of motion in the standard form, it is compatible with any standard integration scheme, though in this submission we limit ourselves to
Implicit Euler.

@@

We derive the mass matrix for our system from the definition of kinetic energy:
\begin{equation}
    T = \frac{1}{2}\dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
\end{equation}
% \begin{equation}
% T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
% \text{,}
% \end{equation}
% where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain. Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
% \begin{equation}
% \begin{split}
% T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
% \end{split}
% \text{.}
% \end{equation}
% The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.

@@

\subsection{External Forces}
We compute the force of gravity and apply external forces in the standard way.
Gravity is defined as $\vc{f}_g = \int_{\Omega} \rho\vc{g} d\Omega$, where $\vc{g}$ is the acceleration due to gravity acting on the object. 
We leverage our meshless quadrature~(\refsec{quadrature}) for integration.

@@

The total error at the boundary of the object can be written as
\begin{equation}
\epsilon = \int_{\partial\Omega}||\Gamma\left(\refX\right)\vc{q} - \Jnurbs\left(\refX\right)\vc{q}||_2^2 d\refX, \\
\end{equation} which is a quadratic.

### 2021-01-27 paper/sections/results.tex
12 pages

\begin{table*}[h]
  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
  }
  \label{tbl:perf}
  \begin{center}
  \begin{tabular}{l c c c c c c c c c c}
   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
   \hline 
   \rowcolor[HTML]{DAE8FC} 
   \textbf{Cantilever (soft)}       & ? & Jelly & $1 \times 10^3$ & $1 \times 10^6$& 0.45 & ? & ? & ? & ? & ? \\
   \textbf{Cantilever (stiff)}      & ? & Rubber & $1 \times 10^3$ & $5 \times 10^6$ & 0.45 & ? & ? & ? & ? & ? \\
   \textbf{Coffee mug}              & ? & Jelly & $1 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ?  \\
   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 \times 10^3$ & $7 \times 10^6$ & 0.4 & ? & ? & ? & ? & ? \\
   \textbf{Rocket (stiff)}          & 648 & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
   \textbf{Castle}                  & ? & Jelly & $1.27 \times 10^3$ & $2 \times 10^3$ & 0.45 & ? & ? & ? & ? & ? \\
   \textbf{Chicken}                 & ? & Rubber & $1.27 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ? \\
   \textbf{Starship}                & ? & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
   \textbf{Tire}                    & ? & Rubber & $2 \times 10^3$ & $1 \times 10^6$ & 0.47 & ? & ? & ? & ? & ? \\
   \hline
  \end{tabular}
  \end{center}

@@

%\dave{we should report which models we made ourselves}
We implemented SEM in a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
respectively. Benchmarks were performed using a MacBook Pro with an Intel i5 2.3GHz processor, 16GB of RAM and an Intel Iris Plus Graphics 655 GPU.

@@

For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and triangulate the NURBS surfaces, which only represents an implementation simplification. 
Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.

@@

% relaxed modelling constraints
By virtue of its meshless nature, SEM is robust to a wide range of challenging models with large gaps and disconnected primitives.
\reffig{fig:chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
Despite the lack of explicit connectivity, the SEM blending weights have the effect of implicitly enforcing connectivity at these seams.

@@

\begin{figure}[H]
  \includegraphics[width=\columnwidth]{figures/mug_overlap}
  \caption{Our SEM is robust to overlapping regions and self-intersections. Here the simulation result of the coffee mug with overlapping handle (bottom) still matches its corresponding counterpart without negligible overlaps (top). }
  % \caption{Coffee mug with overlapping handle. (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}

>>>>

\begin{table*}[h]
  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio. Here the time per step doesn't include the time for collision handling. 
  }
  \label{tbl:perf}
  \begin{center}
  \begin{tabular}{l c c c c c c c c c c}
   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
   \hline 
   \rowcolor[HTML]{DAE8FC} 
   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e10^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
   \textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
   \textbf{Coffee mug}              & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\
   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.4 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ \\
   \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ \\
   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
   \textbf{Chicken}                 & 1773 & Rubber & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
   \textbf{Astronauts}              & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Tire}                    & ? & Rubber/Steel & $2 e^3$ / $3 e^3$ & $1 e^6$ / $7 e^9$ & 0.47 / 0.35 & ? & ? & ? & ? & ? \\
   \hline
  \end{tabular}
  \end{center}

@@

%\dave{we should report which models we made ourselves}
We implemented SEM using a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
respectively. Benchmarks were performed using a MacBook Pro with an Intel i5 2.3GHz processor, 16GB of RAM and an Intel Iris Plus Graphics 655 GPU.

@@

For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and Blender~\cite{blender}, and triangulate the NURBS surfaces, which only represents an implementation simplification. 
Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer that directly accepts NURBS geometry.

@@

% relaxed modelling constraints
By virtue of its meshless nature, SEM is robust to a wide range of challenging models with large gaps and disconnected primitives.
\reffig{chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
Despite the lack of explicit connectivity, the SEM blending weights have the effect of implicitly enforcing connectivity at these seams.

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/mug_overlap}
  \caption{Our SEM is robust to overlapping regions and self-intersections. Here the simulation result of the coffee mug with overlapping handle (bottom) still matches its corresponding counterpart without negligible overlaps (top). }
  % \caption{Coffee mug with overlapping handle. (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}
